---
title: "Final Project: Minecraft: Containers vs. VM | Authors: Liao, Athena; Sharma, Nikki; Franke, Maximilian"
output: html_notebook
---
# Clean workspace
```{r}
# Clean workspace
rm(list = ls())
```
# Install packages
```{r}
#install.packages("ggplot2")
#install.packages("grid")
#install.packages("dplyr")
#install.packages("xtable")
#install.packages("corrgram")
#install.packages("corrplot")
#install.packages("devtools")
#install.packages("broom")
```
# Load packages
```{r}
library(ggplot2)
library(grid)
library(dplyr)
library(xtable)
library(corrgram)
library(corrplot)
#library(readr) 
#library(tidyverse,warn.conflicts = FALSE)
#library(RColorBrewer)
#library(plotly)
options(warn=-1,message=-1)
#library(rcompanion)
library(devtools)
library(broom)
```
# Preparation
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

ggplotRegression <- function (fit) {
    
    require(ggplot2)
    
    ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
        geom_point() +
        stat_smooth(method = "lm", col = "red") +
        labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                           "Intercept =",signif(fit$coef[[1]],5 ),
                           " Slope =",signif(fit$coef[[2]], 5),
                           " P =",signif(summary(fit)$coef[2,4], 5)))
}
```
# Set working directory
```{r}
# set your personal working directory concerning the files
setwd("~/Desktop/05_Big Data Analytics/CIS-541 Fundament. Big Data/06_Final_Project/Data/Final Data")
```
# Load data
```{r}
serverdata <- read.csv("Native.csv", stringsAsFactors = F)
serverdocker <- read.csv("NContainer.csv", stringsAsFactors = F)
VMdata <- read.csv("VM.csv", stringsAsFactors = F)
VMDocker <- read.csv("VMDocker_0910.csv", stringsAsFactors = F)
```
# Bind the df to one df
```{r}
serverdata$case <- "Native"
serverdocker$case <- "Container"
VMdata$case <- "VirtualMachine"
VMDocker$case <-"VM Docker"
# Change column name to bind the datasets
names(serverdata)[names(serverdata) == "Clients"] <- "Client"
names(VMDocker)[names(VMDocker) == "Clients"] <- "Client"
names(VMdata)[names(VMdata) == "Clients"] <- "Client"
names(serverdocker)[names(serverdocker) == "Clients"] <- "Client"

df <- rbind(serverdata, serverdocker, VMdata,VMDocker)
```
# Usage of CPU and Memory in the 3 cases
```{r}
p1 <- df %>% ggplot(aes(factor(case), HostCPU.)) + geom_col(aes(fill=case)) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+
  labs(x = "Case", y = "Host CPU usage", title = "Usage of CPU in 4 cases") 

p2 <- df %>% ggplot(aes(factor(case), HostMem.)) + geom_col(aes(fill=case)) + 
  labs(x = "Case", y = "Host Memory usage", title = "Usage of Memory in 4 cases") +theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+theme(legend.position="none")

layout <- matrix(c(1,1,1,2,2),1,5,byrow=TRUE)
multiplot(p1, p2, layout=layout)
```

# Plot the received & sent Bytes & Packets
```{r}
p1 <- df %>% ggplot(aes(factor(case), Bytes_Sent/10000)) + geom_col(aes(fill=case)) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = "Case", y = "Bytes Sent")

p2 <- df %>% ggplot(aes(factor(case), Bytes_Recv/10000)) + geom_col(aes(fill=case)) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = "Case", y = "Bytes Received") +
  theme(legend.position="none")

p3 <- df %>% ggplot(aes(factor(case), Packets_Sent/10000)) + geom_col(aes(fill=case)) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = "Case", y = "Packets Sent")

p4 <- df %>% ggplot(aes(factor(case), Packets_Recv/10000)) + geom_col(aes(fill=case)) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = "Case", y = "Packets Received") +
  theme(legend.position="none")

layout <- matrix(c(1,1,1,2,2,3,3,3,4,4),2,5,byrow=TRUE)
multiplot(p1, p2, p3, p4, layout=layout)
```
# Plot development of HostCPU concerning Time and the 4 cases
```{r}
# Vmdata
VMdata$min <- substr(VMdata$Time, 1, 5)
VMTime <- select(VMdata, min, HostCPU., Client)
VMTime <- VMTime %>% group_by(Client, min) %>% 
  summarize(average = mean(HostCPU.))
ggplot(data = VMTime, aes(x = min, y = average, color = Client)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + labs(title = "Average Usage per minute of capacity [Virtual Machine]", y = "Average Usage of capacity", x = "Minute")
# VMDocker
VMDocker$min <- substr(VMDocker$Time, 1, 5)
VMDockerTime <- select(VMDocker, min, HostCPU., Client)
VMDockerTime <- VMDockerTime %>% group_by(Client, min) %>% 
  summarize(average = mean(HostCPU.))
ggplot(data = VMDockerTime, aes(x = min, y = average, color = Client)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + labs(title = "Average Usage per minute of capacity [VM Docker]", y = "Average Usage of capacity", x = "Minute")
# serverdata
serverdata$min <- substr(serverdata$Time, 1, 5)
serverdataTime <- select(serverdata, min, HostCPU., Client)
serverdataTime <- serverdataTime %>% group_by(Client, min) %>% 
  summarize(average = mean(HostCPU.))
ggplot(data = serverdataTime, aes(x = min, y = average, color = Client)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + labs(title = "Average Usage per minute of capacity [Native]", y = "Average Usage of capacity", x = "Minute")
# serverdocker
serverdocker$min <- substr(serverdocker$Time, 1, 5)
serverdockerTime <- select(serverdocker, min, HostCPU., Client)
serverdockerTime <- serverdockerTime %>% group_by(Client, min) %>% 
  summarize(average = mean(HostCPU.))
ggplot(data = serverdockerTime, aes(x = min, y = average, color = Client)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + labs(title = "Average Usage per minute of capacity [Container]", y = "Average Usage of capacity", x = "Minute")

```
# Plot development of HostCPU concerning Clients and the 4 cases
```{r}
mt <- ggplot(df, aes(x=Client, y=HostCPU.)) + geom_point(shape=1) + stat_smooth(method = "lm") 
mt + facet_grid(~case)
```
# Plot development of HostCPU, colored by the factor(HostCPU) concerning Clients and the 4 cases
```{r}
p <- ggplot(data = df, aes(x = Client,y= HostCPU., colour = factor(HostCPU.))) + geom_point()
p + facet_grid(~case)
```
# Plot development of HostCPU concerning the Clients
```{r}
df$HostCPU.1<-as.numeric(df$HostCPU.)
df$Clients1<-as.numeric(df$Client)
ggplot(df, aes(Client, HostCPU., col=case, group=1)) + geom_col()+
 facet_grid(case~., scale='free_y') + geom_step(data=df, aes(Clients1, HostCPU.1)) +
  scale_color_discrete(breaks=c('psavert','uempmed','unemploy','q')) + labs(title = "Development of Host CPU concerning the Clients and cases")
```
# Plot HostMem concerning the Clients and the 4 cases
```{r}
df$HostMem.1<-as.numeric(df$HostMem.)
df$Clients1<-as.numeric(df$Clients)
ggplot(df, aes(Client, HostMem., col=case, group=1)) + geom_col()+
 facet_grid(case~., scale='free_y') + geom_step(data=df, aes(Clients1, HostMem.1)) +
  scale_color_discrete(breaks=c('psavert','uempmed','unemploy','q')) + labs(title = "Development of Host Mem concerning the Clients and cases")
```
# Boxplot HostCPU concerning Clients and the 4 cases
```{r}
p <- ggplot(data = df, aes(x = Client, y= HostCPU., fill = case)) + geom_boxplot()
p
```
# Plot HostMem. concerning Clients and colored by the 4 cases
```{r}
p1 <- ggplot(data = df, aes(x = Client, y= HostMem., colour = factor(case))) + geom_point()
p1
```
# Safe the settings for df
```{r}
serverdata$case <- "Native"
serverdocker$case <- "Container"
VMdata$case <- "VirtualMachine"
VMDocker$case <- "Docker"

df <- rbind(serverdata, serverdocker, VMdata, VMDocker)
```
# Plot HostCPU concerning Clients and 4 cases
```{r}
p1 <- aggregate(HostCPU. ~ Client + case, data = df, FUN = mean)
ggplot(data = p1, aes(x = Client, y = HostCPU., color = case)) + geom_point() + geom_line() +
  labs(title = "Usage of CPU in 4 cases", y = "Host CPU") 
```
# Plot HostMem concerning Clients and 4 cases
```{r}
p2 <- aggregate(HostMem. ~ Client + case, data = df, FUN = mean)
ggplot(data = p2, aes(Client, HostMem.,color=case)) + geom_line() + geom_point() + 
  labs(title = "Usage of Memory in 4 cases", y = "Host Memory") + theme(legend.position="none")
```
# Plot Bytes Sent concerning Clients and 4 cases
```{r}
p3 <- aggregate(Bytes_Sent ~ Client + case, data = df, FUN = mean)
ggplot(data = p3, aes(Client, Bytes_Sent,color=case)) + geom_line() + geom_point()  + 
  labs(title = "Bytes sent in 4 cases", y = "Bytes Sent")
```
# Plot Bytes Received concerning Clients and 4 cases
```{r}
p4 <- aggregate(Bytes_Recv ~ Client + case, data = df, FUN = mean)
ggplot(data = p4, aes(Client, Bytes_Recv,color=case)) + geom_line() + geom_point() + 
  labs(title = "Bytes received in 4 cases", y = "Bytes Received")
```
# Plot Packets Sent concerning Clients and 4 cases
```{r}
p5 <- aggregate(Packets_Sent ~ Client + case, data = df, FUN = mean)
ggplot(data = p5, aes(Client, Packets_Sent,color=case)) + geom_line() + geom_point() + 
  labs(title = "Packets sent in 4 cases" , y = " Packets sent ")
```
# Plot Packets Received concerning Clients and 4 cases
```{r}
p6 <- aggregate(Packets_Recv ~ Client + case, data = df, FUN = mean)
ggplot(data = p6, aes(Client, Packets_Recv, color=case)) + geom_line() + geom_point() + 
  labs(title = "Packets received in 4 cases",  y = " Packets received ") 
```
# Correlation for Server Data
```{r}
# select numerical columns for correlation matrix
serverdataNumbers <- select(serverdata, HostCPU., HostMem., Bytes_Sent, Bytes_Recv, Packets_Sent, Packets_Recv, Client)

# corrgram
corrgram(serverdataNumbers)

# Creating a Correlation Matrix of Metric Variables
cor <- cor(serverdataNumbers, use = "pairwise")
cor

# corrplot
corrplot(cor, method = "number")
```
# Correlation for Server Docker
```{r}
# select numerical columns for correlation matrix
serverdockerNumbers <- select(serverdocker, HostCPU., HostMem., Bytes_Sent, Bytes_Recv, Packets_Sent, Packets_Recv, Client)

# corrgram
corrgram(serverdockerNumbers)

# Creating a Correlation Matrix of Metric Variables
cor <- cor(serverdockerNumbers, use = "pairwise")
cor

# corrplot
corrplot(cor, method = "number")
```
# Correlation for Virtual Machine
```{r}
# select numerical columns for correlation matrix
VMNumbers <- select(VMdata, HostCPU., HostMem., Bytes_Sent, Bytes_Recv, Packets_Sent, Packets_Recv, Client)

# corrgram
corrgram(VMNumbers)

# Creating a Correlation Matrix of Metric Variables
cor <- cor(VMNumbers, use = "pairwise")
cor

# corrplot
corrplot(cor, method = "number")
```
# Correlation for Virtual Machine Docker
```{r}
# select numerical columns for correlation matrix
VMDockerNumbers <- select(VMDocker, HostCPU., HostMem., Bytes_Sent, Bytes_Recv, Packets_Sent, Packets_Recv, Client)

# corrgram
corrgram(VMDockerNumbers)

# Creating a Correlation Matrix of Metric Variables
cor <- cor(VMDockerNumbers, use = "pairwise")
cor

# corrplot
corrplot(cor, method = "number")
```
# Correlation for binded df
```{r}
# select numerical columns for correlation matrix
dfNumbers <- select(df, HostCPU., HostMem., Bytes_Sent, Bytes_Recv, Packets_Sent, Packets_Recv, Client)

# corrgram
corrgram(dfNumbers)

# Creating a Correlation Matrix of Metric Variables
cor <- cor(dfNumbers, use = "pairwise")
cor

# corrplot
corrplot(cor, method = "number")
```
# Backward elimination for linear regression model
```{r}

full.model <- lm(Client ~ HostCPU. + HostMem. + Bytes_Sent + Bytes_Recv + Packets_Sent + Packets_Recv, 
                 data = df)
# stepwise
step(object = full.model, # start at the full model
    direction = "backward") # allow it remove predictors but not add them

# the result of the best regression model it could find:
lmBackward <- lm(formula = Client ~ HostCPU. + HostMem. + Bytes_Sent + Bytes_Recv + Packets_Sent + Packets_Recv, 
                data = df)
summary(lmBackward)
```
#Modelling
```{r}
df$case <- as.factor(df$case)
df$Date <- NULL
df$Time <- NULL

names(df)[1:2] <- c("HostCPU", "HostMem")
```
# We model a regression without coefficient in order to get coefficent for all the categories for categorical variable (case).
# CPU usage
```{r}
model_CPU <- lm(HostCPU ~ Client + case -1, data = df)
#CPU_coef <- summary(model_CPU)$coefficients
print(summary(model_CPU))

aov_CPU <- aov(HostCPU ~ Client + case -1, data = df)
summary(aov_CPU)
TukeyHSD(aov_CPU, which = "case")

# Drop Docker, because p-value is high and therefore it should be exluded from this analysis
df3case <- rbind(serverdata, serverdocker, VMdata)
model_CPU2 <- lm(HostCPU. ~ Client + case -1, data = df3case)
print(summary(model_CPU2))
```
# Interpretation
```{r}
# The case docker shows a p-value (0.858), which is higher than 0.05. That means if you want to increase the amount of players with a good stability (a valid measure of stability is CPU usage), the case Docker is not a good solution in comparison to the other three cases.
# If you remove the case Docker, the R-squared increases a bit. That means the "quality" of the model is better. Nevertheless, we want to compare all cases concerning the best case for the maximum number of players.
```
# Memory usage
```{r}
model_mem <- lm(HostMem ~ Client + case -1, data = df)
print(summary(model_mem))
```
# Plotting the fit

# Bytes_Sent
```{r}
model_Bytes_Sent <- lm(Bytes_Sent ~ Client + case -1, data = df)
print(summary(model_Bytes_Sent))
```
# Bytes Received
```{r}
model_Bytes_Recv <- lm(Bytes_Recv ~ Client + case -1, data = df)
print(summary(model_Bytes_Recv))
```
# Packets Sent
```{r}
model_Packets_Sent <- lm(Packets_Sent ~ Client + case -1, data = df)
print(summary(model_Packets_Sent))
```
# Packets Received
```{r}
model_Packets_Recv <- lm(Packets_Recv ~ Client + case -1, data = df)
print(summary(model_Packets_Recv))
```

# Export the lm regression models
```{r}
# model_CPU
tidy_model_CPU <- tidy(model_CPU)
write.csv(tidy_model_CPU, "tidy_model_CPU.csv")
# Model_MEM
tidy_model_mem <- tidy(model_mem)
write.csv(tidy_model_mem, "tidy_model_mem.csv")
# Model_Bytes_Sent
tidy_model_Bytes_Sent <- tidy(model_Bytes_Sent)
write.csv(tidy_model_Bytes_Sent, "tidy_model_Bytes_Sent.csv")
# Model_Bytes_Recv
tidy_model_Bytes_Recv <- tidy(model_Bytes_Recv)
write.csv(tidy_model_Bytes_Recv, "tidy_model_Bytes_Recv.csv")
# Packets_Sent
tidy_model_Packets_Sent <- tidy(model_Packets_Sent)
write.csv(tidy_model_Packets_Sent, "tidy_model_Packets_Sent.csv")
# Packets_Recv
tidy_model_Packets_Recv <- tidy(model_Packets_Recv)
write.csv(tidy_model_Packets_Recv, "tidy_model_Packets_Recv.csv")
# lmbackward
tidy_model_backward <- tidy(lmBackward)
write.csv(tidy_model_backward, "tidy_model_backward.csv")
```
# Hypothesis Testing
# Claim: The mean of Host CPU is significantly higher after 40 clients than under 40 clients
```{r}
df$CPUCat <- if_else(df$Client >= 40, "CPU with >= 40 Clients", "CPU with < 40 Clients")
wilcox.test(df$HostCPU ~ df$CPUCat, mu=0, alt = "less", conf.int = TRUE, conf.level = 0.95, paired = FALSE)
```
# Claim: The mean of Host Memory is significantly higher after 40 clients than under 40 clients
```{r}
df$MEMCat <- if_else(df$Client >= 40, "MEM with >= 40 Clients", "MEM with < 40 Clients")
wilcox.test(df$HostMem ~ df$MEMCat, mu=0, alt = "less", conf.int = TRUE, conf.level = 0.95, paired = FALSE)
```
# Claim: The mean of Bytes sent is significantly higher after 40 clients than under 40 clients
```{r}
df$BytesSentCat <- if_else(df$Client >= 40, "MEM with >= 40 Clients", "MEM with < 40 Clients")
wilcox.test(df$Bytes_Sent ~ df$BytesSentCat, mu=0, alt = "less", conf.int = TRUE, conf.level = 0.95, paired = FALSE)
```
# Claim: The mean of Mem Bytes received is significantly higher after 40 clients than under 40 clients
```{r}
df$BytesRecvCat <- if_else(df$Client >= 40, "MEM with >= 40 Clients", "MEM with < 40 Clients")
wilcox.test(df$Bytes_Recv ~ df$BytesRecvCat, mu=0, alt = "less", conf.int = TRUE, conf.level = 0.95, paired = FALSE)
```
# Interpretation
```{r}
# As proved with the Wilcox.Test, there is a significantly difference concerning valid measures of stability (e.g. Usage Host CPU or Usage Host MEM) after 40 Clients and before. After 40 Clients the performance decreases.
```
# Is there a sig. difference between the averages per minutes (Usage Host CPU) concerning the 4 cases
```{r}
serverdataTime$case <- "Native"
serverdockerTime$case <- "Container"
VMTime$case <- "VirtualMachine"
VMDockerTime$case <-"VM Docker"
# Bind the df
Time <- rbind(serverdataTime, serverdockerTime, VMDockerTime, VMTime)
# anova
aov_Time <- aov(Time$average ~ Client + case + min, data = Time)
summary(aov_Time)
```
# Interpretation
```{r}
# The mean averages per minutes are different concerning the different Clients, cases and minutes.
```

